---
title: "Peer Review Assignment 1: Due on Sunday, April 7 at midnight"
author: enter your name here
format: html
editor: visual
---

## Grading Instructions

You will receive 60 points for submitting the assignment, full points for each correct part as shown below and no points for incorrect answers.

For graders: Please download an assigned student's file, type your comments and corrections, enter the grade on Camino and upload back the file by Tuesday, April 9 at midnight.

------------------------------------------------------------------------

## Getting data from FRED: (10 points)

For this assignment you need to access FRED database and get Real GDP per Capita for the US (ID# A939RX0Q048SBEA).

Read the following instructions to learn how to obtain access to FRED database using FRED API. <https://cran.r-project.org/web/packages/fredr/vignettes/fredr.html>

Once you get access to FRED, retrieve Real GDP per Capita between the first quarter of 1970 and the last quarter of 2023.

Show your code below and make sure to comment your code.

Do not display your FRED API key, it is a private information.

```{r}
#| label: setup
#| echo: true
#| message: true

```

Now that you have retrieved the data, answer the following questions:

## What is the type of this data? (6 points)

```{r}
#| echo: true
#| message: true


```

## Coerce your data frame to tsibble: (6 points)

```{r}
#| echo: true
#| message: true


```

## Change the series_id variable values to "RGDP per capita": (6 points)

```{r}
#| echo: true
#| message: true


```

## What was RGDP per capita in the first quarter of 2021?: (6 points)

```{r}
#| echo: true
#| message: true


```

## What was an annualized rate of RGDP per capita growth over the time period of your data? : (6 points)

```{r}
#| echo: true
#| message: true


```
