---
title: "Time Series Analysis: Week 2 - April 8"
author: Orkhan Hasanaliyev
format: 
  revealjs:
    transition: slide
    slide-number: true
editor: visual
---

## 

*"You can observe a lot just by watching"*\
                                                                                                          Yogi Berra

## Load ffp3 package

-   Load the following packages

```{r}
#| label: setup
#| echo: true
#| message: false
library(tidyverse)
library(fpp3)
library(tswge)

```

## Graphs

Plots allow us to identify:

-   Patterns

-   Unusual observations

-   Changes over time

-   Relationships between variables.

## US GDP

Filter the US data from global_economy

```{r}
#| echo: true
#| message: true
us_economy <- global_economy %>%
filter(Country == "United States")

print(us_economy)
```

## The US GDP

Plot the US GDP

```{r}
#| echo: true
#| message: true
us_economy |> autoplot(GDP) +
    labs(title = "Nominal GDP, the US ",
       subtitle = "",
       y = "Nominal GDP (trln US$)")

```

## Global Temperature

```{r}
#| echo: true
#| message: true

global_temp<- tibble(
  Year = 1850:2009,
  Temperature = hadley
)

global_temp<- global_temp|>
  as_tsibble(index=Year)


global_temp|>
  autoplot(Temperature)

```

## Time series patterns

Trend

:   pattern exists when there is a long-term increase or decrease in the data.

Seasonal

:   pattern exists when a series is influenced by seasonal factors (e.g., the quarter of the year, the month, or day of the week).

Cyclic

:   pattern exists when data exhibit rises and falls that are *not of fixed period*

## Time series components

Differences between seasonal and cyclic patterns:

-   seasonal pattern constant length; cyclic pattern variable length
-   average length of cycle longer than length of seasonal pattern
-   magnitude of cycle more variable than magnitude of seasonal pattern

## Domestic Car Production

```{r}
#| echo: true
#| message: true

library(fredr)

fredr_set_key("33223ed79fd61ea6915872b9a2cc2256")

car_prod <- fredr(
  series_id = "DAUPNSA"
)

car_prod<- car_prod|>
  mutate(date = yearmonth(date)) |>
  as_tsibble(
    index = date
  )

```

## Domestic Car Production

```{r}
#| echo: true
#| message: true

car_prod|>autoplot()
```

## Domestic Car Production

Use `gg_season()` to plot the data against the individual "seasons"

```{r}
#| echo: true
#| message: true

car_prod|>
  gg_season()
```

## Australian Electricity Production

```{r}
#| echo: true
#| message: true
aus_production |>
  filter(year(Quarter) >= 1980) |>
  autoplot(Electricity) +
  labs(y = "GWh", title = "Australian electricity production")
```

## The US Retail Employment

```{r}
#| echo: true
#| message: true
us_employment |>
  filter(Title == "Retail Trade", year(Month) >= 1980) |>
  autoplot(Employed / 1e3) + geom_point() + 
  labs(y = "Million people", title = "Retail employment, USA")
```

## The US Retail Employment

```{r}
#| echo: true
#| message: true
us_employment |>
  filter(Title == "Retail Trade", year(Month) >= 1980) |>
  gg_season()+
  labs(y = "Million people", title = "Retail employment, USA")
```

## US Gas Consumption

```{r}
#| echo: true
#| message: true

#install.packages('USgas')
library(USgas)

us_total

```

## CA and ME gas consumption

```{r}
#| echo: true
#| message: true

us_tsibble <- us_total |>
  as_tsibble(index=year, key=state)
# Draw CA and ME gas consumption over time
us_tsibble |>
  filter(state %in% c("California","Maine")) |>
  autoplot(y/1e3) +
  labs(y = "billion cubic feet")

```

## Unemployment Claims

```{r}
#| echo: true
#| message: true

un_claims <- fredr(
  series_id = "CCNSA"
)

un_claims<- un_claims|>
  mutate(date = yearweek(date)) |>
  as_tsibble(
    index = date
  )
```

## Unemployment Claims

```{r}
#| echo: true
#| message: true

un_claims|>autoplot()+
    labs(title = "Continued Claims ",
       subtitle = "Insured Unemployment",
       y = "Number")
```

## Unemployment Claims

```{r}
#| echo: true
#| message: true
un_claims|>
  gg_season()
```

## Electricity Consumption

```{r}
#| echo: true
#| message: true
vic_elec 

```

## Electricity Consumption

```{r}
#| echo: true
#| message: true
vic_elec |> autoplot()

```

## Electricity Consumption

```{r}
#| echo: true
#| message: true
vic_elec |> gg_season(Demand)

```

## Electricity Consumption

```{r}
#| echo: true
#| message: true
vic_elec |> gg_season(Demand, period = "week")

```

## Electricity Consumption

```{r}
#| echo: true
#| message: true
vic_elec |> gg_season(Demand, period = "day")

```

## The US Retail Employment

```{r}
#| echo: true
#| message: true
us_employment |>
  filter(Title == "Retail Trade", year(Month) >= 1980) |>
  gg_subseries()+
  labs(y = "Million people", title = "Retail employment, USA" )+
  aes(color=Title)
```

## The US Retail Employment

```{r}
#| echo: true
#| message: true
us_employment |>
  filter(Title %in% c("Retail Trade", "Mining and Logging", "Professional and Business Services", "Education and Health Services"), year(Month) >= 1980) |>
  autoplot()+
  labs(y = "Million people", title = "Employment by Sector, USA")+   aes(color=Title)
```

## The US Retail Employment

```{r}
#| echo: true
#| message: true
us_employment |>
  filter(Title %in% c("Retail Trade", "Mining and Logging", "Professional and Business Services", "Education and Health Services"), year(Month) >= 1980) |>
  gg_season()+
  labs(y = "Million people", title = "Employment by Sectors, USA")
```

## The US Retail Employment

```{r}
#| echo: true
#| message: true
us_employment |>
  filter(Title %in% c("Retail Trade", "Mining and Logging", "Professional and Business Services", "Education and Health Services"), year(Month) >= 1980) |>
  gg_subseries(period="month")+
  facet_wrap(vars(Title), nrow = 2, scales = "free_y")+
  labs(y = "Million people", title = "Employment by Sectors, USA")
```
